---
title: "Appendix: R Code for TRAIN Law Analysis"
author: "Jannisah M. Palo"
format: pdf
editor: source
warning: false
---

# Appendix: R Code

## A.1 Data Loading

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(readr)
library(data.table)
library(dplyr)
library(ggplot2)
library(scales)
library(broom)
library(fixest)
library(EValue)

# Load raw FIES volumes (two-volume structure per year)
v1_2012 <- fread("fies_2012_vol.1.csv")
v2_2012 <- fread("fies_2012_vol.2.csv")

v1_2015 <- fread("fies_2015_vol.1.csv")
v2_2015 <- fread("fies_2015_vol.2.csv")

v1_2018 <- fread("fies_2018_vol.1.csv")
v2_2018 <- fread("fies_2018_vol.2.csv")
```

---

## A.2 Merging Volumes and Deduplication Checks

### 2012

```{r}
keys_2012 <- c("W_REGN", "W_OID", "W_SHSN", "W_HCN")

# Check for duplicates in each volume
nrow(v1_2012[, .N, by = keys_2012][N > 1])
nrow(v2_2012[, .N, by = keys_2012][N > 1])

# Merge volumes: drop overlapping non-key columns from vol.2
setkeyv(v1_2012, keys_2012)
setkeyv(v2_2012, keys_2012)
v2_2012_clean <- v2_2012[, !setdiff(intersect(names(v1_2012), names(v2_2012)), keys_2012),
                          with = FALSE]
fies_2012 <- v1_2012[v2_2012_clean]

# Validate merge
nrow(v1_2012); nrow(fies_2012)
fies_2012[, .N, by = keys_2012][N > 1]
all.equal(v1_2012$RFACT, fies_2012$RFACT)
```

### 2015

```{r}
keys_2015 <- c("W_REGN", "W_ID", "W_SHSN", "W_HCN")

nrow(v1_2015[, .N, by = keys_2015][N > 1])
nrow(v2_2015[, .N, by = keys_2015][N > 1])

setkeyv(v1_2015, keys_2015)
setkeyv(v2_2015, keys_2015)
v2_2015_clean <- v2_2015[, !setdiff(intersect(names(v1_2015), names(v2_2015)), keys_2015),
                          with = FALSE]
fies_2015 <- v1_2015[v2_2015_clean]

nrow(v1_2015); nrow(fies_2015)
fies_2015[, .N, by = keys_2015][N > 1]
all.equal(v1_2015$RFACT, fies_2015$RFACT)
```

### 2018

```{r}
keys_2018 <- c("W_REGN", "W_PROV", "SEQUENCE_NO")

nrow(v1_2018[, .N, by = keys_2018][N > 1])
nrow(v2_2018[, .N, by = keys_2018][N > 1])

setkeyv(v1_2018, keys_2018)
setkeyv(v2_2018, keys_2018)
v2_2018_clean <- v2_2018[, !setdiff(intersect(names(v1_2018), names(v2_2018)), keys_2018),
                          with = FALSE]
fies_2018 <- v1_2018[v2_2018_clean]

nrow(v1_2018); nrow(fies_2018)
fies_2018[, .N, by = keys_2018][N > 1]
all.equal(v1_2018$RFACT, fies_2018$RFACT)
```

---

## A.3 Region 10 Datasets and Stacking

```{r}
# Load pre-processed Region 10 files
fies_12 <- read_csv("fies_2012_combined.csv") %>%
  rename(TOTEX = T_TOTEX, FOOD = T_FOOD,
         NFOOD = T_NFOOD, TRANSPORT = T_TRANSPORT) %>%
  mutate(year = 2012)

fies_15 <- read_csv("fies_2015_combined.csv") %>%
  mutate(post = 0, year = 2015)

fies_18 <- read_csv("fies_2018_combined.csv") %>%
  mutate(post = 1, year = 2018)

# Stack all three years
combined <- bind_rows(fies_12, fies_15, fies_18)
```

---

## A.4 Variable Construction

### Treatment Variable

```{r}
# Transport exposure thresholds are defined using 2015 baseline distribution
base_2015 <- combined %>%
  filter(year == 2015)

q1 <- quantile(base_2015$TRANSPORT, 0.10, na.rm = TRUE)  # low-exposure cutoff
q3 <- quantile(base_2015$TRANSPORT, 0.90, na.rm = TRUE)  # high-exposure cutoff

# treat: Binary indicator based on 2015 transport expenditure percentiles
# 1 = high exposure (top 10%); 0 = low exposure (bottom 10%)
# Middle 80% excluded from DiD sample
```

### Post-Intervention Indicator

```{r}
# post: Binary time indicator
# 0 = 2015 (pre-TRAIN Law); 1 = 2018 (post-TRAIN Law)
```

### Outcome Variable

```{r}
# log_totex: Natural log of total household expenditure (TOTEX)
# Log transformation applied to reduce skewness and allow percentage interpretation
```

### Control Variables

```{r}
# FSIZE : Household size
# AGE   : Age of household head
# URB   : Urban/rural indicator (1 = urban, 0 = rural)
# RFACT : Survey expansion weight (used as regression weight)
```

### Construct DiD Dataset

```{r}
did_data <- combined %>%
  filter(year %in% c(2015, 2018)) %>%
  mutate(
    treat = case_when(
      TRANSPORT >= q3 ~ 1,
      TRANSPORT <= q1 ~ 0,
      TRUE            ~ NA_real_
    ),
    post      = if_else(year == 2018, 1, 0),
    log_totex = log(TOTEX)
  ) %>%
  filter(!is.na(treat))

table(did_data$year)
table(did_data$treat, did_data$post)
```

---

## A.5 Difference-in-Differences Model

```{r}
did_model <- lm(
  log_totex ~ treat * post + FSIZE + AGE + URB,
  data    = did_data,
  weights = RFACT
)
summary(did_model)
```

---

## A.6 DiD Plot

```{r}
# Weighted group means
trend_data <- did_data %>%
  filter(!is.na(log_totex), !is.na(treat), !is.na(RFACT)) %>%
  group_by(year, treat) %>%
  summarise(mean_log_totex = weighted.mean(log_totex, w = RFACT, na.rm = TRUE),
            .groups = "drop")

# Counterfactual for treated group
low_2015  <- trend_data %>% filter(year == 2015, treat == 0) %>% pull(mean_log_totex)
low_2018  <- trend_data %>% filter(year == 2018, treat == 0) %>% pull(mean_log_totex)
high_2015 <- trend_data %>% filter(year == 2015, treat == 1) %>% pull(mean_log_totex)
high_2018 <- trend_data %>% filter(year == 2018, treat == 1) %>% pull(mean_log_totex)

cf_high_2018 <- high_2015 + (low_2018 - low_2015)

cf_line <- data.frame(
  year           = c(2015, 2018),
  mean_log_totex = c(high_2015, cf_high_2018)
)
```

```{r, echo=FALSE}
train_did_plot <- ggplot(
  trend_data,
  aes(x = year, y = mean_log_totex,
      color = factor(treat), group = factor(treat))
) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 3) +
  geom_line(data = cf_line,
            aes(x = year, y = mean_log_totex),
            inherit.aes = FALSE,
            linetype = "dashed", color = "black", linewidth = 1) +
  geom_segment(x = 2018, xend = 2018,
               y = high_2018, yend = cf_high_2018,
               color = "#FF69B4", linewidth = 1.2) +
  scale_color_manual(values = c("0" = "#E07B6A", "1" = "#5B9BD5"),
                     labels = c("Low exposure", "High exposure")) +
  scale_x_continuous(breaks = c(2012, 2015, 2018)) +
  labs(
    title = "Log Total Expenditure Trends of High and Low Exposure Groups",
    x     = "Year",
    y     = "Weighted Mean log Total Expenditure",
    color = "Group"
  ) +
  theme_minimal(base_size = 15) +
  theme(
    plot.title         = element_text(face = "bold", size = 15, hjust = 0),
    axis.title         = element_text(size = 15, face = "bold"),
    axis.text          = element_text(size = 15),
    legend.title       = element_text(face = "bold", size = 15),
    legend.text        = element_text(size = 15),
    panel.grid.major   = element_line(color = "gray90", linewidth = 0.4),
    panel.grid.minor   = element_blank(),
    plot.margin        = margin(15, 20, 15, 15)
  )

ggsave("did_plot_train_final.png", plot = train_did_plot,
       width = 8, height = 5, dpi = 300, bg = "white")
train_did_plot
```

---

## A.7 Parallel Trends Assumption Tests

### A.7.1 Pre-Trend Regression Test (2012–2015)

```{r}
# Apply 2015-defined thresholds to 2012 and 2015 data
pretrend_data <- combined %>%
  filter(year %in% c(2012, 2015)) %>%
  mutate(
    treat = case_when(
      TRANSPORT >= q3 ~ 1,
      TRANSPORT <= q1 ~ 0,
      TRUE            ~ NA_real_
    ),
    log_totex = log(TOTEX),
    pre       = if_else(year == 2015, 1, 0)
  ) %>%
  filter(!is.na(treat))

table(pretrend_data$year)
table(pretrend_data$treat, pretrend_data$year)
```

```{r}
pretrend_model <- lm(
  log_totex ~ treat * pre + FSIZE + AGE + URB,
  data    = pretrend_data,
  weights = RFACT
)
summary(pretrend_model)
```

### Pre-Trend Plot

```{r, echo=FALSE}
pretrend_plot_data <- pretrend_data %>%
  group_by(year, treat) %>%
  summarise(mean_log_totex = weighted.mean(log_totex, RFACT, na.rm = TRUE),
            .groups = "drop") %>%
  mutate(group = if_else(treat == 1,
                         "High transport exposure",
                         "Low transport exposure"))

train_pretrend_plot <- ggplot(pretrend_plot_data,
       aes(x = year, y = mean_log_totex,
           color = group, group = group)) +
  geom_line(linewidth = 1.2) +
  geom_point(size = 3) +
  scale_x_continuous(breaks = c(2012, 2015)) +
  labs(
    title = "Pre-Trend Check (2012\u20132015)",
    x     = "Year",
    y     = "Weighted Mean of Log Total Expenditure",
    color = "Group"
  ) +
  theme_minimal(base_size = 15) +
  theme(
    plot.title       = element_text(face = "bold", size = 15, hjust = 0),
    axis.title       = element_text(size = 15, face = "bold"),
    axis.text        = element_text(size = 15),
    legend.title     = element_text(face = "bold", size = 15),
    legend.text      = element_text(size = 15),
    panel.grid.major = element_line(color = "gray90", linewidth = 0.4),
    panel.grid.minor = element_blank(),
    plot.margin      = margin(15, 20, 15, 15)
  )

ggsave("pretrend_plot_train.png", plot = train_pretrend_plot,
       width = 8, height = 5, dpi = 300, bg = "white")
train_pretrend_plot
```

### A.7.2 Placebo Test

```{r}
# Placebo DiD using only pre-treatment years (2012 vs. 2015)
# A significant interaction here would undermine parallel trends
placebo_model <- lm(
  log_totex ~ treat * pre + FSIZE + AGE + URB,
  data    = pretrend_data,
  weights = RFACT
)
summary(placebo_model)
```

---

## A.8 Event Study

```{r}
# All three survey years; 2015 as reference year (last pre-treatment period)
combined_es <- combined %>%
  filter(year %in% c(2012, 2015, 2018)) %>%
  mutate(
    treat     = case_when(
      TRANSPORT >= q3 ~ 1,
      TRANSPORT <= q1 ~ 0,
      TRUE            ~ NA_real_
    ),
    log_totex = log(TOTEX),
    year      = relevel(factor(year), ref = "2015")
  ) %>%
  filter(!is.na(treat))

event_model <- lm(
  log_totex ~ treat * year + FSIZE + AGE + URB,
  data    = combined_es,
  weights = RFACT
)
summary(event_model)
```

### Event Study Plot

```{r, echo=FALSE}
event_results <- tidy(event_model) %>%
  filter(grepl("treat:year", term)) %>%
  mutate(
    year    = c(2012, 2018),
    ci_low  = estimate - 1.96 * std.error,
    ci_high = estimate + 1.96 * std.error
  )

train_event_study_plot <- ggplot(event_results,
       aes(x = year, y = estimate)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray30") +
  geom_errorbar(aes(ymin = ci_low, ymax = ci_high),
                width = 0.4, linewidth = 0.9) +
  geom_point(size = 3) +
  # Reference year point
  annotate("point", x = 2015, y = 0, size = 3, shape = 21,
           fill = "white", color = "gray50", stroke = 1.2) +
  annotate("text", x = 2015, y = -0.005,
           label = "Reference\n(2015)", vjust = 1.4,
           size = 3.5, color = "gray50", fontface = "italic") +
  scale_x_continuous(breaks = c(2012, 2015, 2018)) +
  labs(
    title = "Event Study: Effect of TRAIN Law on Total Expenditure",
    x     = "Year",
    y     = "Treatment Effect on log Total Expenditure"
  ) +
  theme_minimal(base_size = 15) +
  theme(
    plot.title       = element_text(face = "bold", size = 15, hjust = 0),
    axis.title       = element_text(size = 15, face = "bold"),
    axis.text        = element_text(size = 15),
    panel.grid.major = element_line(color = "gray90", linewidth = 0.4),
    panel.grid.minor = element_blank(),
    plot.margin      = margin(15, 20, 15, 15)
  )

ggsave("event_study_train.png", plot = train_event_study_plot,
       width = 8, height = 5, dpi = 300, bg = "white")
train_event_study_plot
```



# =============================================================================
# E-VALUE SENSITIVITY ANALYSIS FOR DiD
# =============================================================================

## E-Value Input Parameters

```{r}
# Extract DiD coefficient and standard error from the DiD model
est  <- coef(did_model)["treat:post"]
print(est)
se <- coef(summary(did_model))["treat:post", "Std. Error"]
print(se)
sd_y <- sigma(did_model)
print(sd_y)
```

## E-Value Results

```{r}
EValue::evalues.OLS(est = est, se = se, sd = sd_y, delta = 1)
```

## E-Value Contour Plot

```{r}
# ── 1. TRAIN DiD model ───────────────────────────────────────────────────────
did_model <- lm(
  log_totex ~ treat * post + FSIZE + AGE + URB,
  data = did_data,
  weights = RFACT
)

# ── 2. Extract DiD coefficient, SE, and residual SD ──────────────────────────
est    <- coef(did_model)["treat:post"]
se   <- coef(summary(did_model))["treat:post", "Std. Error"]
sd_y <- sigma(did_model)

# ── 3. Compute E-values ──────────────────────────────────────────────────────
ev_train <- EValue::evalues.OLS(
  est   = est,
  se    = se,
  sd    = sd_y,
  delta = 1
)

# ── 4. Extract RR and E-values ───────────────────────────────────────────────
rr_point <- ev_train["RR", "point"]
rr_lower <- ev_train["RR", "lower"]
rr_upper <- ev_train["RR", "upper"]

e_point <- ev_train["E-values", "point"]
e_lower <- ev_train["E-values", "lower"]
e_upper <- ev_train["E-values", "upper"]

# Because the TRAIN estimate is negative, RR < 1.
# For plotting, use the reciprocal RR so the E-value contour is on RR > 1 scale.
rr_point_plot <- ifelse(rr_point < 1, 1 / rr_point, rr_point)

# For a negative/protective estimate, the CI limit closest to the null is the upper RR.
rr_ci_plot <- ifelse(rr_upper < 1, 1 / rr_upper, rr_upper)

# For negative estimates, the relevant CI E-value is in the "upper" column.
e_ci <- e_upper

# If CI crosses the null, E-value for CI becomes 1.
if (is.na(e_ci)) e_ci <- 1

# ── 5. Correct E-value contour function ──────────────────────────────────────
evalue_curve <- function(x, rr) {
  rr * (x - 1) / (x - rr)
}

# ── 6. Curve data ────────────────────────────────────────────────────────────
df_point <- data.frame(
  x = seq(rr_point_plot + 0.0001, 1.40, length.out = 1000)
)

df_point$y <- evalue_curve(df_point$x, rr_point_plot)

df_point <- subset(df_point, y >= 1.00 & y <= 1.40)

# CI curve
df_ci <- data.frame(
  x = seq(rr_ci_plot + 0.0001, 1.40, length.out = 1000)
)

df_ci$y <- evalue_curve(df_ci$x, rr_ci_plot)

df_ci <- subset(df_ci, y >= 1.00 & y <= 1.40)

# ── 7. Annotation data frames ────────────────────────────────────────────────
point_label <- data.frame(
  x     = e_point,
  y     = e_point,
  label = sprintf("E-value: (%.3f, %.3f)", e_point, e_point)
)

ci_label <- data.frame(
  x     = e_ci,
  y     = e_ci,
  label = sprintf("E-value (CI): (%.3f, %.3f)", e_ci, e_ci)
)

# ── 8. Plot ──────────────────────────────────────────────────────────────────
p_train <- ggplot() +
  
  # Background
  annotate(
    "rect",
    xmin = 1.00, xmax = 1.40,
    ymin = 1.00, ymax = 1.40,
    fill = "grey80",
    alpha = 0.30
  ) +
  
  # Null reference line
  geom_hline(
    yintercept = 1,
    linetype   = "dashed",
    linewidth  = 0.7,
    color      = "grey50"
  ) +
  
  # Point-estimate E-value curve
  geom_line(
    data = df_point,
    aes(x = x, y = y),
    color = "#3266AD",
    linewidth = 1.3
  ) +
  
  # CI E-value curve
  geom_line(
    data = df_ci,
    aes(x = x, y = y),
    color = "grey55",
    linewidth = 1.0,
    linetype = "dashed"
  ) +
  
  # Point-estimate E-value marker
  geom_point(
    data = point_label,
    aes(x = x, y = y),
    shape = 21,
    size = 3,
    fill = "white",
    color = "#3266AD",
    stroke = 1.6
  ) +
  
  # Point-estimate label
  geom_text(
    data = point_label,
    aes(x = x, y = y, label = label),
    nudge_x = 0.012,
    nudge_y = 0.010,
    hjust = 0,
    size = 3.5,
    color = "#3266AD"
  ) +
  
  # CI E-value marker
  geom_point(
    data = ci_label,
    aes(x = x, y = y),
    shape = 21,
    size = 3,
    fill = "white",
    color = "grey45",
    stroke = 1.6
  ) +
  
  # CI label
  geom_text(
    data = ci_label,
    aes(x = x, y = y, label = label),
    nudge_x = 0.012,
    nudge_y = -0.020,
    hjust = 0,
    size = 3.5,
    color = "grey35"
  ) +
  
  labs(
    title    = "E-value Contour Plot",
    subtitle = "TRAIN Law ATT Estimate",
    x        = "Confounder–Exposure association (RR scale)",
    y        = "Confounder–Outcome association (RR scale)"
  ) +
  
  coord_cartesian(
    xlim = c(1.00, 1.40),
    ylim = c(1.00, 1.40),
    clip = "off"
  ) +
  
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(
      face = "bold",
      size = 14,
      hjust = 0,
      color = "grey10",
      margin = margin(b = 4)
    ),
    plot.subtitle = element_text(
      size = 10,
      hjust = 0,
      color = "grey45",
      margin = margin(b = 10)
    ),
    axis.title = element_text(
      face = "bold",
      size = 11,
      color = "grey20"
    ),
    axis.text = element_text(
      color = "grey30",
      size = 10
    ),
    axis.line = element_line(
      color = "grey30",
      linewidth = 0.4
    ),
    panel.grid.major = element_line(
      color = "grey88",
      linewidth = 0.35
    ),
    panel.grid.minor = element_blank(),
    plot.margin = margin(12, 55, 12, 12)
  )

p_train

# Save plot
ggsave(
  "evalue_did_train_plot.png",
  plot = p_train,
  width = 8,
  height = 5,
  dpi = 300,
  bg = "white"
)
```




